\chapter{Bregman divergences}

This chapter studies a concept which generalizes Euclidean distance. This
generalization measures the discrepancy between two points using the curvature
of a given convex function.   

\section{Definition and properties}

\index{Legendre function}
Let $\phi : \R^d \to (-\infty,\infty]$ be a function of Legendre type, which we
recall (Chapter \ref{sec:conjugate_smoothness}) means that $\phi$ is essentially
smooth and essentially strictly convex. The \emph{Bregman divergence} $D_\phi$
corresponding to $\phi$ is a map defined for \smash{$x \in \bar\Omega =
  \dom(\phi)$} and $y \in \Omega = \interior(\dom(\phi))$ by     
\index{Bregman divergence}
\begin{equation}
\label{eq:bregman_divergence}
D_\phi(x,y) = \phi(x) - \phi(y) - \nabla\phi(y)^\T (x - y).
\end{equation}
The value $D_\phi(x,y)$ can be interpreted as the gap between the value
$\phi(x)$ and its first-order approximation $\phi(y) + \nabla\phi(y)^\T (x -
y)$ centered at $y$. Since $\phi$ is convex, we know that $D_\phi(x,y) \geq
0$. Further, since $\phi$ is strictly convex on the line segment joining $x$ and 
$y$, we have $D_\phi(x,y) > 0$ for $x \not= y$. 

In what follows, we refer to $\phi$ as the \emph{potential function} underlying
the Bregman divergence $D_\phi$. An important special case is the quadratic
potential \smash{$\phi(x) = \frac{1}{2} \|x\|_2^2$}. In this case, one can check
that \smash{$D_\phi(x,y) = \frac{1}{2} \|x-y\|_2^2$}, half of the squared
Euclidean distance between $x$ and $y$. In general, we note that $D_\phi$ is not
necessarily a proper distance (or monotone transformation of a distance); for
example, it does not even need to be symmetric in its arguments. Still, Bregman 
divergences admit a number of useful properties, described next.    

\paragraph{Convex in its first argument.}
\parlab{par:bregman_first}

For any fixed $y$, the map $x \mapsto D_\phi(x,y)$ is convex. To see this, note
that this map can be written as $x \mapsto \phi(x) + a^\T x + b$, where $a,b$ 
depend on $y$. This means that a Bregman ball centered around the second
argument, of the form 
\[
\{ x : D_\phi(x,y) \leq r \}
\]
for fixed $y,r$, is a convex set. Furthermore, Bregman balls of this type are
also compact if $\phi$ is closed (Exercise \ref{ex:bregman_first}).     

\paragraph{Star-shaped in its second argument.} 
\parlab{par:bregman_second}

Meanwhile, for fixed $x$, the map $y \mapsto D_\phi(x,y)$ is not necessarily
convex. However, it satisfies a weaker property: it is \emph{star-shaped}, with
the point $x$ called the star-center. Precisely, this means that $t \mapsto
D_\phi(x, (1-t) x + ty)$ is nonincreasing on $[0,1]$ for any point $y$
(Exercise \ref{ex:bregman_second}). Consequently, a Bregman ball centered
around the first argument,   
\[
\{ y : D_\phi(x,y) \leq r \}
\]
for fixed $x,r$, is what is known as a \emph{star-convex} set: for any $y$ in
this set, the line segment joining $x$ and $y$ lies entirely in the set. 
\index{star-convex set}

\paragraph{Conditions for quadratic growth.}

If $\phi$ is Lipschitz smooth with parameter $L>0$, then it is easy to verify
(using Theorem \ref{thm:lipschitz_smoothness} part (iii)) that 
\index{Bregman divergence!quadratic growth}
\[
D_\phi(x,y) \leq \frac{L}{2} \|x-y\|_2^2.
\]
If $\phi$ is strongly convex with parameter $m>0$, then it is similarly easy to
verify (using Theorem \ref{thm:strong_convexity} part (iii)) that 
\[
D_\phi(x,y) \geq \frac{m}{2} \|x-y\|_2^2.
\]
Therefore, for a Lipschitz smooth and strongly convex potential $\phi$, the
Bregman divergence $D_\phi(x,y)$ grows quadratically in $x-y$ (in the sense that
it is upper and lower bounded by constant multiples of the squared Euclidean
distance between $x$ and $y$).

\paragraph{Three-point identity.}
\parlab{par:bregman_three_point}

Euclidean distance satisfies the following \emph{three-point identity}: 
\[
\frac{1}{2} \|x-z\|_2^2 = \frac{1}{2} \|x-y\|_2^2 + \frac{1}{2} \|y-z\|_2^2 + 
(y-z)^\T (x-y), 
\]
for any points $x,y,z \in \R^d$, which can be checked by direct
calculation. This is sometimes called the \emph{law of cosines}. To interpret 
this, note that the third term on the right-hand side is $(y-z)^\T (x-y) =
-\|x-y\|_2 \|z-y\|_2 \cos(\theta)$, where $\theta$ is the angle in between $x-y$  
and $z-y$; thus the above identity relates one side length of a triangle with
vertices $x,y,z$ to the other two side lengths and the cosine of the angle in
between them. (This generalizes the Pythagorean theorem for right triangles,
which corresponds to $\cos(\theta) = 0$). Bregman divergences admit an
analogous three-point identity:     
\index{Bregman divergence!three-point identity}
\[
D_\phi(x,z) = D_\phi(x,y) + D_\phi(y,z) + (\nabla\phi(y) - \nabla(\phi(z))^\T
(x-y),  
\]
for any \smash{$x \in \bar\Omega$} and $y,z \in \Omega$, which can again be
checked by direct calculation.  
\medskip

\begin{Example}
A key example of a Bregman divergence is the Kullback-Leibler divergence. This
corresponds to the negative entropy potential function   
\[
\phi(x) = \sum_{i=1}^d x_i \log(x_i),
\]
also called the negative Shannon entropy. This is Legendre with $\dom(\phi) = \{
x : \one^\T x = 1, \, x \geq 0 \}$, the probability simplex. In this case, we
write $D_\phi(x,y) = \KL(x,y)$, where
\index{Kullback-Leibler divergence} 
\begin{equation}
\label{eq:kullback_leibler}
\KL(x,y) = \sum_{i=1}^d x_i \log(x_i / y_i),
\end{equation}
verified by direct calculation. The Kullback-Leibler divergence is a notable 
case (like squared Euclidean distance) where the Bregman divergence is jointly
convex in $x,y$ (Exercise \ref{ex:loss_functions} part d), recalling that in
general we only know it is convex in $x$ for fixed $y$.
\end{Example}

\section{Bregman projection}

% In what ways do these preserve properties of projections/proximals? In what
% ways do they not?

% This paper: https://cmps-people.ok.ubc.ca/bauschke/Research/103.pdf may help
% as a resource 

Pythagorean theorem from law of cosines

\section{Conjugacy connections}

% For any Legendre function h and its conjugate h^*:
% D_h(y, x) = D_{h^*}(u, v)
% with u = \nabla h(x) and v = \nabla h(y). Note the reversal of the order of
% arguments! Proof is elementary

\section{Exponential families*}

% In exponential families, the gradient of the LPF is the map between natural
% parameter and the mean (cf Barndoff and Nielsen, Brad, Martin and Mike).  

% Maybe these notes help as as resource (aside from those books):
% https://sites.stat.washington.edu/mmp/courses/stat538/winter12/Handouts/l7-exponential.pdf 

% The inverse, the map from the mean to the natural parameter, is the gradient
% of the conjugate! 

% Deviance in exponential family can be written in terms of a Bregman
% divergence, in TWO ways. These two ways are related by conjugacy.

% The usual "primal" representation D_\phi(\theta,\theta^*) is most useful for
% training since this loss is convex in \theta. The alternative "dual"
% representation D_\phi(y,\mu) is more useful for evaluation b/c the covariance
% formula for excess risk works more cleanly this way. Brad's work shows this. 

\section{Proper loss functions*}

% Connection between proper losses and conjugacy (cf Savage, Gneiting and
% Raftery, etc.) 

% Conjugacy relation gives rise to a way to define a surrogate loss (credit to
% John Duchi)

\SkipTocEntry\section*{Chapter notes}

\clearpage

\begin{xcb}{Exercises}
\begin{enumerate}[label=\thechapter.\arabic*]
\settowidth{\leftmargini}{00.00.\hskip\labelsep}
\item Let $f$ be a closed convex function. Prove that $u \in
  \interior(\dom(f^*))$ if and only if $x \mapsto f(x) - u^\T x$ has no
  directions of recession. Hint: use Exercise \ref{ex:conjugate_recession} part
  (i) combined with the translation rule for convex conjugates in Property
  \parref{par:conjugate_translation}.  

\item \label{ex:bregman_first} 
  Let $\phi$ be a closed Legendre potential function, and fix $y \in
  \interior(\dom(\phi))$. 

\begin{enumerate}[label=\alph*.]
\item Use the last exercise to prove that $x \mapsto \phi(x) - \nabla\phi(y)^\T
  x$ has no directions of recession. Hint: use Theorem
  \ref{thm:conjugate_legendre} to show that $\nabla\phi(y) \in
  \interior(\dom(\phi^*))$. 

\item Use part a to prove that $x \mapsto D_\phi(x,y)$ has no directions of
  recession, and therefore has compact sublevel sets. Hint: recall Lemma
  \ref{lem:direction_recession}. 
\end{enumerate}

\item \label{ex:bregman_second}
  Let $\phi$ be a Legendre potential function, and fix $x \in
  \interior(\dom(\phi))$.

\begin{enumerate}[label=\alph*.]
\item Fix $v \in \R^d$, and define $d_x(t) = D_\phi(x, x+tv)$ for $t \in \R$.
  For $t \geq s \geq 0$, check that   
  \[
  d_x(t) - d_x(s) = \phi(x+sv) - \phi(x+tv) + (t\nabla(x+tv) - s\nabla(x+sv))^\T
  v. 
  \]

\item Using convexity of $\phi$, show that 
  \[
  d_x(t) - d_x(s) \geq s (\nabla(x+tv) - \nabla(x+sv))^\T v \geq 0. 
  \]
  Use this to verify the star-shaped claim in Property
  \parref{par:bregman_second}.   

\item Now suppose $D_\phi(x,y) \leq r$. Use the star-shaped property just
  established to verify the star-convexity claim in Property
  \parref{par:bregman_second}, that is, $D_\phi(x, (1-t) x + ty) \leq r$ for $t
  \in [0,1]$.   
\end{enumerate}
\end{enumerate}
\end{xcb}