\chapter{Duality}
\label{chap:duality}

\section{LP duality*}
\label{sec:lp_duality}

Duality is fascinating topic in mathematical optimization: the basic arguments
used in the theory of duality are elementary and yet they can lead to powerful
and far-reaching conclusions.    

The story begins, for us, with linear programs (LPs). Consider a generic LP of
the form 
\begin{equation}
\label{eq:lp_primal}
\begin{alignedat}{2}
&\minimize_x \quad && c^\T x \\
&\st && Ax \leq b \\
& && Gx = h, 
\end{alignedat}
\end{equation}
where $c \in \R^d$, $A \in \R^{m \times d}$, $b \in \R^m$, $G \in \R^{k \times 
  d}$, $h \in \R^k$. The reason we start the chapter by studying LPs is that,   
in this problem class, we can build up dual problems ``constructively''; this 
constructive approach is not possible for general optimization problems, and
helps us appreciate the importance and elegance of Lagrange duality, which is
covered next.     

The fundamental question that underlies the study of duality is as follows: what
is the tightest lower bound we can form on the optimal criterion value $f^\star
= c^\T x^\star$ in \eqref{eq:lp_primal}? To address this, suppose that $u \in
\R^m$ and $v \in \R^k$ are arbitrary vectors---which we call \emph{dual 
  variables} in this context---with $u$ nonnegative in each component, $u \geq   
0$. Provided that $x \in \R^d$ is a feasible point for problem
\eqref{eq:lp_primal}, it holds that $u^\T (Ax - b) \leq 0$ and $v^\T (Gx - h) =
0$, and so, adding these together gives        
\begin{equation}
\label{eq:lp_dual_nonnegative}
u^\T (Ax - b) + v^\T (Gx - h) \leq 0.
\end{equation}
In order to obtain a lower bound on the criterion value $c^\T x$, we rearrange
the above into
\[
(-A^\T u - G^\T v)^\T x \geq -b^\T u - h^\T v.
\]
The key observation is that this provides the lower bound we desire provided
that the dual variables $u,v$ are chosen such that $-A^\T u - G^\T v= c$. This
is true for any feasible $x$, thus taking an infimum over all such $x$ gives   
\[
f^\star \geq -b^\T u - h^\T v, \quad \text{for any $u,v$ such that $-A^\T u -
  G^\T v = c$ and $u \geq 0$}.
\]
Finally, to make this lower bound as tight as possible, we maximize the
right-hand side above,  
\begin{equation}
\label{eq:lp_weak_duality}
f^\star \geq \, \underbrace{\sup \big\{ -b^\T u - h^\T v :  A^\T u + G^\T v =
  -c, \, u \geq 0 \big\}}_{g^\star}.
\end{equation}
Now the right-hand side in \eqref{eq:lp_weak_duality}, which we may denote as 
$g^\star = -b^\T u^\star - h^\T v^\star$, is itself the optimal criterion value
associated with an optimization problem, indeed an LP,     
\index{dual problem!linear program}
\begin{equation}
\label{eq:lp_dual}
\begin{alignedat}{2}
&\maximize_{u,v} \quad && b^\T u - h^\T v \\
&\st && A^\T u + G^\T v = -c \\
& && u \geq 0.
\end{alignedat}
\end{equation}
In this context, we call \eqref{eq:lp_dual} the \emph{dual LP} and the original 
problem \eqref{eq:lp_primal} the \emph{primal LP}. Note that, by construction 
\eqref{eq:lp_weak_duality}, we have $f^\star \geq g^\star$: the optimal
criterion in the dual problem is a lower bound on the optimal criterion in the  
primal problem.  

There are several natural follow-up questions that we can ask: for example, when
does equality hold in \eqref{eq:lp_weak_duality}, $f^\star = g^\star$? And how
do solutions $x^\star$ and $u^\star, v^\star$ in \eqref{eq:lp_primal} and
\eqref{eq:lp_dual}, respectively, relate? We will address these questions and
more, over this chapter and the next one. First, however, we must develop a
theory of duality beyond LPs. 

\begin{Remark}
In the attempt to move beyond LPs with the constructive approach to duality, we 
run into a shortcoming: when the criterion $f$ is nonlinear, but the constraint
functions are linear, we have no way in general to combine the
constraints---which are linear equalities and inequalities in $x$---to construct
a lower bound on $f(x)$. But there is another way: Lagrangian duality, as we
will see next, applies seamlessly to LPs and convex optimization more broadly
(and even some nonconvex problems).   
\end{Remark}

\section{Lagrangian duality}
\label{sec:lagrangian_duality}

Lagrangian duality (or Lagrange duality) starts with the same motivation as in
the last section, but cast in a more general setting: we seek a lower bound on
the optimal criterion value $f^\star$ in           
\begin{equation}
\label{eq:primal_problem}
\begin{alignedat}{2}
&\minimize_x \quad && f(x) \\
&\st \quad && h_i(x) \leq 0, \; i=1,\ldots,m \\ 
& && \ell_j(x) = 0, \; j=1,\ldots,k.
\end{alignedat}
\end{equation}
At the moment, we do not assume that criterion $f$ or the constraint functions
$h_i$, $i=1,\dots,m$, and $\ell_j$, $j=1,\dots,k$ are convex. Thus, to be clear,
we do not assume that \eqref{eq:primal_problem} is a convex problem. 

As before, let $u \in \R^m$ and $v \in \R^k$ be arbitrary vectors, with $u \geq
0$, which we call \emph{dual variables} in the current context. Using the same
basic idea as in \eqref{eq:lp_dual_nonnegative}, note that for any feasible $x$
for \eqref{eq:primal_problem}, 
\[
\sum_{i=1}^m u_i h_i(x) + \sum_{j=1}^r v_j \ell_j(x) \leq 0.
\]
We now add this quantity to the criterion $f(x)$ to define what we call the
\emph{Lagrangian function},   
\index{Lagrangian function}
\[
L(x,u,v) = f(x) + \sum_{i=1}^m u_i h_i(x) + \sum_{j=1}^r v_j \ell_j(x).
\] 
By construction, this provides a lower bound on the criterion,
\begin{equation}
\label{eq:lagrangian_lb1}
f(x) \geq L(x,u,v), \quad \text{for any feasible $x$ and $u,v$ such that $u
  \geq 0$},
\end{equation}
and minimizing over the feasible set, denoted $C \subseteq \R^d$, yields
\begin{equation}
\label{eq:lagrangian_lb2}
f^\star \geq \inf_{x \in C} \, L(x,u,v) \geq \inf_x \, L(x,u,v), \quad \text{for
  any $u,v$ such that $u \geq 0$}.  
\end{equation}

\begin{Remark}
The second inequality in \eqref{eq:lagrangian_lb2} is really the key idea behind
Lagrangian duality; had we stopped at the first inequality in
\eqref{eq:lagrangian_lb2}, we would have not have (yet) gotten anywhere
practically interesting, because minimizing the Lagrangian over $x \in C$ is 
hard, in general---no easier than minimizing $f$ over $x \in C$, which is our
original primal problem \eqref{eq:primal_problem}. Meanwhile, minimizing the
Lagrangian over $x \in \R^d$ is usually more tractable (as we will see in
examples that follow), and still produces an effective lower bound.
\end{Remark}

It is convenient to denote the right-most quantity in \eqref{eq:lagrangian_lb2}
by 
\index{dual function}
\begin{equation}
\label{eq:dual_function}
g(u,v) = \inf_x \, L(x,u,v), 
\end{equation}
which we call the \emph{Lagrange dual function}. Then \eqref{eq:lagrangian_lb2}
reads $f^\star \geq g(u,v)$ for any $u,v$ such that $u \geq 0$, and so to
obtain the tightest lower bound, we maximize $g$, which yields
\index{dual problem}
\begin{equation}
\label{eq:dual_problem}
\maximize_{u,v} \quad g(u,v) \quad \st \quad u \geq 0,
\end{equation}
which we call the \emph{Lagrange dual problem} associated with
\eqref{eq:primal_problem}. Our convention will be to drop the term ``Lagrange''
when referring to the dual function and dual problem defined above.

\section{Properties and examples}

Two key properties of Lagrange duality, which follow more or less immediately
from the construction in the last section, are described next.

\paragraph{Weak duality.}
\parlab{par:weak_duality}



\paragraph{Convexity of dual problem.}
\parlab{par:convex_dual}

hi

\medskip

\begin{Example}
LPs.
\end{Example}

\begin{Example}
QPs.
\end{Example}

\section{Interpretations}
\label{sec:duality_interpretations}

\section{SDP duality*}
\label{sec:sdp_duality}

% for SDP duality: take a look at 
% https://www.mit.edu/~parrilo/cdc03_workshop/ejc03_comp.pdf
% https://ocw.mit.edu/courses/electrical-engineering-and-computer-science/6-251j-introduction-to-mathematical-programming-fall-2009/readings/MIT6_251JF09_SDP.pdf